---
title: 'Report'
author: 'Sang Doan'
date: '10/15/2020'
output: pdf_document
---

```{r setup, include = F}
source('config.R')
```
# Problem 1
See `data_munge.R`.

# Problem 2
See `data.R`.

# Problem 3
```{r}
pop <- getPop()
facts <- getFacts()
deaths <- getDeaths()
head(get_counties(pop))
```

# Problem 4
```{r}
dem1 <- 'AGE775214'
head(map_counties(deaths, pop, facts, dem1))
```

# Problem 5
Five demographics I chose are:\
  (1) AGE775214 Persons 65 years and over, percent, 2014\
  (2) EDU685213 Bachelor's degree or higher, percent of persons age 25+, 2009-2013\
  (3) RHI225214 Black or African American alone, percent, 2014\
  (4) RHI325214 American Indian and Alaska Native alone, percent, 2014\
  (5) RHI125214 White alone, percent, 2014\
  
## People 65 years and over
```{r}
plotDem(deaths, pop, facts, 'AGE775214') %>% format(scientific = F)
```
\pagebreak

## People with a Bachelor's degree or higher
```{r}
plotDem(deaths, pop, facts, 'EDU685213') %>% format(scientific = F)
```
\pagebreak

## Black or African Americans
```{r}
plotDem(deaths, pop, facts, 'RHI225214') %>% format(scientific = F)
```
\pagebreak

## American Indians and Alaska Natives
```{r}
plotDem(deaths, pop, facts, 'RHI325214') %>% format(scientific = F)
```
\pagebreak

## Whites
```{r}
plotDem(deaths, pop, facts, 'RHI125214') %>% format(scientific = F)
```

Old age is the best predictor of Covid-19 deaths among the five demographics, but all correlations assessed above are insignificant. 

\pagebreak

# Code
`data_munge.R`
```{r eval = F, echo = T}
deaths <- read.csv('../raw_data/covid_deaths.csv', header=T, stringsAsFactors = F)
facts <- read.csv('../raw_data/county_facts.csv', header=T, stringsAsFactors = F)
pop <- read.csv('../raw_data/county_population.csv', header=T, stringsAsFactors = F)

names(deaths)[1:3] <- c('FIPS', 'county', 'state')
names(facts)[1:3] <- c('FIPS', 'county', 'state')
names(pop)[1:3] <- c('FIPS', 'county', 'state')

write.csv(deaths, '../processed_data/processed_covid_deaths.csv', row.names = F)
write.csv(facts, '../processed_data/processed_county_facts.csv', row.names = F)
write.csv(pop, '../processed_data/processed_county_population.csv', row.names = F)
```

`data.R`
```{r eval = F, echo = T}
getFacts <- function() {
  return(read.csv('../processed_data/processed_county_facts.csv', header = T,
                  stringsAsFactors = F))
}

getDeaths <- function() {
  return(read.csv('../processed_data/processed_covid_deaths.csv', header = T,
                  stringsAsFactors = F))
}

getPop <- function() {
  return(read.csv('../processed_data/processed_county_population.csv', 
                  header = T, stringsAsFactors = F))
}
```

`analysis.R`
```{r eval = F, echo = T}
get_counties <- function(d) {
  return(dplyr::select(d, county, state))
}

map_counties <- function(cd, cp, cf, demographic) {
  
  #It appears that all data frames are already arranged in an ascending order by FIPS.
  #In such case, the following three lines are unnecessary.
  cd <- cd %>% arrange(FIPS)
  cf <- cf %>% arrange(FIPS)
  cp <- cp %>% arrange(FIPS)
  
  counties <- get_counties(cd)
  counties$deathRate <- rowSums(select(cd, -FIPS, -county, -state, -stateFIPS)) / cp$population
  counties$statistic <- cf[[demographic]]
  
  return(counties)
}

plotDem <- function(cd, cp, cf, dem) {
  thisDem <- select(map_counties(cd, cp, cf, dem), deathRate, statistic)

  thisPlot <- ggplot(data = thisDem, aes(x = statistic, y = deathRate)) +
    ggtitle(paste('Relationship between Covid-19 Death Rate \nand Percentage of', dem,
                  'in Population by Counties')) +
    geom_point(color = 'red', size = 1.5, fill = 'blue', shape = 1, stroke = 0.5) + 
    xlab(paste('Percentage of', dem, 'in Population')) +
    geom_smooth(method = 'lm', formula = y ~ ifelse(is.finite(log(x)), log(x), 0), 
                se = F, size = 0.5) +
    scale_y_continuous(name = 'Death Rate', limits = c(0, 0.005)) +
    theme_minimal()
  
  print(thisPlot)
  
  return(coef(lm(deathRate ~ statistic, thisDem))[[2]])
  
}
```

`config.R`
```{r eval = F, echo = T}
source('data.R')
source('analysis.R')

library(tidyverse)
library(magrittr)
```